---
title: "Exploratory Data Analysis for NSQIP"
author: "Vidya Koesmahargyo"
date: "11/8/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing libraries
```{r, include=FALSE}
library(dplyr)
library(readr)
library(data.table)
library(measurements)
library('scales')
library(ggplot2)
```

Importing Dataset (already selected columns)
```{r}
newdata <- fread("NSQIP.csv", sep="auto")

```

Dataset structure
```{r}
nsqip <- newdata
str(nsqip)

```

Look at non-cardiac surgeries only.
```{r}
#select out cardiac surgery
nsqip <- filter(nsqip, SURGSPEC != "Cardiac Surgery")
colnames <- colnames(nsqip)

```

Change thecharacter variables into factors for plotting/analysis purposes
```{r, include=FALSE}
#factorize character variables
for (column in colnames) {
  if (class(nsqip[,(column)]) == "character") {
    nsqip[,(column)] <- as.factor(nsqip[,(column)])
    print(class(nsqip[,(column)]))
  }
}
str(nsqip)
```

Examine how much of the data set is missing, explicitlylt (NAs) or implicitly (-99).
```{r}
#find percent NAs
na_lst <- sapply(nsqip, function(x) sum(is.na(x)))
na_lst
percent_na <- lapply(na_lst, function(x) percent((x/nrow(nsqip))))
percent_na

```

```{r}
#implicit  NAs/ = '-99' list for each column
implicit_lst <- sapply(nsqip, function(x) sum(x == -99))
implicit_lst

```

```{r}
#percent implicit for each column (made NAs, since they were null values in the beginning)
implicit_lst[is.na(implicit_lst)] <- 0
percent_implicit <- sapply(implicit_lst, function(x) percent(x/nrow(newdata)))
percent_implicit

#find all missing values
missing_lst <- implicit_lst + na_lst
missing_lst
percent_missing <- lapply(missing_lst, function(x) percent((x/nrow(nsqip))))
percent_missing
```


BMI is not available in dataset, mutate to combine height/weight

```{r}
#combine height and weight to make BMI
nsqip$HEIGHT <- conv_unit(nsqip$HEIGHT, from = "inch", to= "m")
nsqip$WEIGHT <- conv_unit(nsqip$WEIGHT,from ="lbs", to = "kg")

nsqip <- mutate(nsqip, BMI = WEIGHT/HEIGHT^2)
summary(nsqip$BMI)
```


```{r}
#remove NAs in MACE
nsqip$DEATH <- cut(nsqip$DOpertoD, c(-100,0,30))
levels(nsqip$DEATH) <- c("Alive", "Within 30 Days")

```

Stratify accordnig to hematocrit
```{r}
# find percent within hematocrit thresholds (# of patients are columns)
HCT1 <- nsqip %>% filter(PRHCT <= 24 & PRHCT >0)
HCT2 <- nsqip %>% filter(PRHCT <= 27 & PRHCT >24)
HCT3 <- nsqip %>% filter(PRHCT <= 30 & PRHCT >27)
HCT4 <- nsqip %>% filter(PRHCT <= 33 & PRHCT >30)
HCT5 <- nsqip %>% filter(PRHCT <= 36 & PRHCT >33)
HCT6 <- nsqip %>% filter(PRHCT <= 39 & PRHCT >36)
HCT7 <- nsqip %>% filter(PRHCT >39)
```


```{r}
# hematocrit stratification, number of patients in each
dim(HCT1)
dim(HCT2)
dim(HCT3)
dim(HCT4)
dim(HCT5)
dim(HCT6)
dim(HCT7)
```


```{r}
#filter patients for target range in preop values

target_BUN <- nsqip %>% filter(PRBUN >= 5 & PRBUN <= 60)
target_CREAT <- nsqip %>% filter(PRCREAT >= 0.02 & PRCREAT <= 15)
target_PLT <- nsqip %>% filter(PRPLATE>= 50 & PRPLATE <= 700)
target_PTT <- nsqip %>% filter(PRPTT >= 5 & PRPTT <= 25)
target_INR <-nsqip %>% filter(PRINR >= 0.1 & PRINR <= 10)

dim(target_BUN)
dim(target_CREAT)
dim(target_PLT)
dim(target_PTT)
dim(target_INR)

```

```{r}
#number of patients within all target ranges, for reference
in_range <- nsqip %>% filter(PRCREAT >= 0.02 & PRCREAT <= 15, PRPLATE>= 50 & PRPLATE <= 700,PRPTT >= 5 & PRPTT <= 25,PRINR >= 0.1 & PRINR <= 10)
dim(in_range)
```

```{r}
# Look at days to MACE within 30 days for all patients
ALLMACE_30 <- nsqip %>% filter(DOpertoD >0, DCNSCVA >0, DCDMI >0)
dim(ALLMACE_30)
CVA_30 <- nsqip %>% filter(DCNSCVA >0)
dim(CVA_30)
DEATH_30<- nsqip %>% filter(DOpertoD >0)
dim(DEATH_30)
MI_30 <-nsqip %>% filter(DCDMI >0)
dim(MI_30)

# Look at # of MACE occurences within 30 days for all patients

```

```{r}
#NA bool mask

# Remove all NAs to make no_NA dataframe

#impute NAs in implicit list as 0, no implicit NAs in dataframe

## could not do these steps because of NAs, might need to impute some
```





```{r}
#for each obs, create new factor variable with either occurence of death, MI, or stroke.
#plot occurence of MIin each HCT category
nsqip$HCTcat<-cut(nsqip$PRHCT, c(0,24,27,30,33,36,39,300))
levels(nsqip$HCTcat) <- c("<24", "24-27","27-30","30-33","33-36","36-39","39<")
summary(nsqip$HCTcat)

#plots on MACE
ggplot(data = nsqip) + 
  geom_bar(mapping = aes(x = HCTcat, fill= CDMI),position = "dodge")
ggplot(data = nsqip) + 
  geom_bar(mapping = aes(x = HCTcat, fill= CNSCVA),position = "dodge")
ggplot(data = nsqip) + 
  geom_bar(mapping = aes(x = HCTcat, fill= CDARREST),position = "dodge")
ggplot(data = nsqip) + 
  geom_bar(mapping = aes(x = HCTcat, fill= DEATH),position = "dodge")
```

